---
title: "PRSmix in COPDGene NHW with training"
output:
  html_document:
    toc: true
    toc_float: true
date: "`r format(Sys.time(), '%B %d, %Y')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE}
library(data.table)
library(tableone)
library(dplyr)
library(PRSmix)
library(MASS)
library(ggplot2)
library(pROC)
library(ResourceSelection)
```

```{r}
options(bitmapType='cairo')

# import functions 
source("PRSmix_functions.R")
```

```{r}
# cgNhw df with PRSs from get_PRS.Rmd
in_path <- "data/pheno_nhw_prs.csv"
df <- fread(in_path, data.table=F)
```


```{r}
prs_list <- names(df)[grep("_pgs$", names(df))]
prs_list <- c(prs_list, "FEV1_ePRS", "FF_ePRS")

# make binary copd variable from spirometry 
df$copd <- ifelse(df$FEV1_FVC_post_P1 < 0.7 & df$fev1_pp_gli_global_P1 < 80, 1, 
                      ifelse(df$FEV1_FVC_post_P1 >= 0.7 & df$fev1_pp_gli_global_P1 >= 80, 0, NA))

outcomes <- c("total_Net_Exacerbations", "lung_density_vnb_P1", "Pi10_Thirona_P1", "copd")

pcs=paste0("PC",1:5)
covars <- c("Age_P1", "gender", "ATS_PackYears_P1", "BMI_P1")

print(length(prs_list))
print(prs_list)
```

```{r}
sapply(df[c(prs_list, pcs, covars)], function(x) sum(is.na(x)))

sapply(df[outcomes], function(x) sum(is.na(x)))
```

```{r}
df$exacerbation_rate = df$total_Net_Exacerbations / df$total_days_since_P1 * 365

df %>% filter(copd == 1) %>% pull(exacerbation_rate) %>% summary()
df %>% filter(copd == 1) %>% pull(exacerbation_rate) %>% sd(na.rm = TRUE)
df %>% filter(copd == 1) %>% pull(exacerbation_rate) %>% IQR(na.rm = TRUE)
```

# get summary table
```{r}

allVars <- c("Age_P1", "gender", "SmokCigNow_P1", "ATS_PackYears_P1", "BMI_P1", "fev1_pp_gli_global_P1", "FEV1_FVC_post_P1", "copd", "exacerbation_rate")
catVars <- c("gender", "SmokCigNow_P1", "copd")
table1 <- CreateTableOne(data = df, vars = allVars, factorVars = catVars)
table1_output <- print(table1, showAllLevels = TRUE)

write.table(table1_output, "data/Table1_demographics_cgNhw.txt", sep = "\t", row.names = TRUE, quote = FALSE)
```

# loop over outcomes and call PRSmix
```{r}
prsmix_covar = c(covars, pcs)
print(prsmix_covar)

cont_outcomes <- c("total_Net_Exacerbations", "lung_density_vnb_P1", "Pi10_Thirona_P1")
binary_outcome <- "copd"
```

```{r results='hide', message=FALSE, warning=FALSE, eval=FALSE}
# continuous outcomes
weights_table_list <- list()

for (outcome in cont_outcomes) {
  call_PRSmix(df, outcome, outcome_binary=FALSE, prsmix_covar, prs_list, coh = "cgNhw", cat_covar_list = c("gender"))
  out <- scale_PRSmix(df, outcome, coh = "cgNhw")
  df <- out$df
  weights_table <- out$weights_table
  
  weights_table_list[[outcome]] <- weights_table
}
```

```{r results='hide', message=FALSE, warning=FALSE, eval=FALSE}
# binary outcomes
call_PRSmix(df, binary_outcome, outcome_binary=TRUE, prsmix_covar, prs_list, coh = "cgNhw", cat_covar_list = c("gender"))
out <- scale_PRSmix(df, binary_outcome, coh = "cgNhw")
df <- out$df
weights_table <- out$weights_table

weights_table_list[[binary_outcome]] <- weights_table
```

```{r eval=FALSE}
all_weights_table <- Reduce(function(x, y) merge(x, y, by = "prs", all = TRUE), weights_table_list)
all_weights_table <- all_weights_table %>% 
  arrange(desc(copd)) %>%
  mutate_if(is.numeric, signif, digits=3)

write.table(all_weights_table, "data/TableS3_PRSmix_weights_cgNhw.txt", sep = "\t", row.names = FALSE, quote = FALSE)
```

```{r}
all_weights_table<-read.table("data/Table2_PRSmix_weights_cgNhw.txt", sep = '\t', header = TRUE)
DT::datatable(all_weights_table)
```

```{r eval=FALSE}
# save the df
fwrite(df, file = "data/cgNhw_PRSmix_out_df.csv")
```

```{r}
in_path <- "data/cgNhw_PRSmix_out_df.csv"
df<-fread(in_path, data.table=F)
```

# Each outcome predicted by its corresponding PRS
```{r message=FALSE, warning=FALSE}
total_Net_Exacerbations_adj <- call_model(df, outcome_name = "total_Net_Exacerbations", pcs, covars, 
                                             model_type = "nb", adjustment = "offset(log(total_days_since_P1/365))", coh="cgNhw")
lung_density_vnb_P1_adj <- call_model(df, outcome_name = "lung_density_vnb_P1", pcs, covars, 
                                         model_type = "lm", adjustment = "scanner_model_clean_P1", coh="cgNhw")
Pi10_Thirona_P1_adj <- call_model(df, outcome_name = "Pi10_Thirona_P1", pcs, covars, 
                                  model_type = "lm", adjustment = "scanner_model_clean_P1", coh="cgNhw")
copd_no_adj <- call_model(df, outcome_name = "copd", pcs, covars, 
                                  model_type = "binomial", adjustment = NULL, coh="cgNhw")
```

```{r}
summary_table <- Reduce(function(x, y) merge(x, y, by = "prs", all = TRUE), 
                        list(total_Net_Exacerbations_adj[, c(1, 5)], 
                             lung_density_vnb_P1_adj[, c(1, 5)],
                             Pi10_Thirona_P1_adj[, c(1, 5)],
                             copd_no_adj[, c(1, 5)]))

DT::datatable(summary_table)
```

# Table3_PRSs_copd: copd predicted by single PRSs and PRSmix
```{r message=FALSE, warning=FALSE}
copd_prs_list <- c("FF_ePRS", "FEV1_ePRS", "Cosentino_pgs", "smoking_pgs", "IPF_pgs", "BMI_pgs", "CRP_pgs", "mix_prs_copd", "prs_composite")

Table3_PRSs_copd <- data.frame()

for (prs in copd_prs_list) {
  model <- build_model(df, "copd", prs, pcs, covars, model_type = "binomial", adjustment=NULL)
  beta <- coef(model)[prs]
  ci <- confint(model, prs, level = 0.95)
  beta_ci <- sprintf("%.3f (%.3f, %.3f)", beta, ci[1], ci[2])
  se <- summary(model)$coefficients[prs, "Std. Error"]
  pval <- summary(model)$coefficients[prs, 4]
  Table3_PRSs_copd <- rbind(Table3_PRSs_copd, 
                                     data.frame(prs = prs, beta = beta, ci_lower = ci[1], ci_upper = ci[2], 
                                                beta_ci = beta_ci, se = se, pval = pval))
}

Table3_PRSs_copd <-Table3_PRSs_copd %>%  mutate_if(is.numeric, signif, digits=3)

DT::datatable(Table3_PRSs_copd, rownames = FALSE)
```

## adjust for baseline FEV1
```{r message=FALSE, warning=FALSE}
Table3_PRSs_copd_adj_fev1 <- data.frame()

for (prs in copd_prs_list) {
  model <- build_model(df, "copd", prs, pcs, covars, model_type = "binomial", adjustment="FEV1_post_P1")
  beta <- coef(model)[prs]
  ci <- confint(model, prs, level = 0.95)
  beta_ci <- sprintf("%.3f (%.3f, %.3f)", beta, ci[1], ci[2])
  se <- summary(model)$coefficients[prs, "Std. Error"]
  pval <- summary(model)$coefficients[prs, 4]
  Table3_PRSs_copd_adj_fev1 <- rbind(Table3_PRSs_copd_adj_fev1, 
                                     data.frame(prs = prs, beta = beta, ci_lower = ci[1], ci_upper = ci[2], 
                                                beta_ci = beta_ci, se = se, pval = pval))
}

Table3_PRSs_copd_adj_fev1 <-Table3_PRSs_copd_adj_fev1 %>%  mutate_if(is.numeric, signif, digits=3)

DT::datatable(Table3_PRSs_copd_adj_fev1, rownames = FALSE)
```

# Table4_PRSs_exacerbations: exacerbation predicted by single PRSs and PRSmix
```{r message=FALSE, warning=FALSE}
Table4_PRSs_exacerbations <- data.frame()

for (prs in copd_prs_list) {
  model <- build_model(df, "total_Net_Exacerbations", prs, pcs, covars, model_type = "nb", adjustment="offset(log(total_days_since_P1/365))")
  beta <- coef(model)[prs]
  ci <- confint(model, prs, level = 0.95)
  beta_ci <- sprintf("%.3f (%.3f, %.3f)", beta, ci[1], ci[2])
  se <- summary(model)$coefficients[prs, "Std. Error"]
  pval <- summary(model)$coefficients[prs, 4]
  Table4_PRSs_exacerbations <- rbind(Table4_PRSs_exacerbations, 
                                     data.frame(prs = prs, beta = beta, ci_lower = ci[1], ci_upper = ci[2], 
                                                beta_ci = beta_ci, se = se, pval = pval))
}

write.table(Table4_PRSs_exacerbations, "data/Table4_PRSs_exacerbations_cgNhw.txt", sep = "\t", row.names = FALSE, quote = FALSE)
Table4_PRSs_exacerbations <-Table4_PRSs_exacerbations %>%  mutate_if(is.numeric, signif, digits=3)
DT::datatable(Table4_PRSs_exacerbations, rownames = FALSE)
```


## adjust for baseline FEV1
```{r message=FALSE, warning=FALSE}
Table4_PRSs_exacerbations_adj_fev1 <- data.frame()

for (prs in copd_prs_list) {
  model <- build_model(df, "total_Net_Exacerbations", prs, pcs, covars, model_type = "nb", adjustment="offset(log(total_days_since_P1/365)) + FEV1_post_P1")
  beta <- coef(model)[prs]
  ci <- confint(model, prs, level = 0.95)
  beta_ci <- sprintf("%.3f (%.3f, %.3f)", beta, ci[1], ci[2])
  se <- summary(model)$coefficients[prs, "Std. Error"]
  pval <- summary(model)$coefficients[prs, 4]
  Table4_PRSs_exacerbations_adj_fev1 <- rbind(Table4_PRSs_exacerbations_adj_fev1, 
                                     data.frame(prs = prs, beta = beta, ci_lower = ci[1], ci_upper = ci[2], 
                                                beta_ci = beta_ci, se = se, pval = pval))
}

Table4_PRSs_exacerbations_adj_fev1 <-Table4_PRSs_exacerbations_adj_fev1 %>%  mutate_if(is.numeric, signif, digits=3)
DT::datatable(Table4_PRSs_exacerbations_adj_fev1, rownames = FALSE)
```

```{r message=FALSE, warning=FALSE}
Table4_PRSs_exacerbations_adj_copd <- data.frame()

for (prs in copd_prs_list) {
  model <- build_model(df, "total_Net_Exacerbations", prs, pcs, covars, model_type = "nb", adjustment="offset(log(total_days_since_P1/365)) + copd")
  beta <- coef(model)[prs]
  ci <- confint(model, prs, level = 0.95)
  beta_ci <- sprintf("%.3f (%.3f, %.3f)", beta, ci[1], ci[2])
  se <- summary(model)$coefficients[prs, "Std. Error"]
  pval <- summary(model)$coefficients[prs, 4]
  Table4_PRSs_exacerbations_adj_copd <- rbind(Table4_PRSs_exacerbations_adj_copd, 
                                     data.frame(prs = prs, beta = beta, ci_lower = ci[1], ci_upper = ci[2], 
                                                beta_ci = beta_ci, se = se, pval = pval))
}

Table4_PRSs_exacerbations_adj_copd <-Table4_PRSs_exacerbations_adj_copd %>%  mutate_if(is.numeric, signif, digits=3)
DT::datatable(Table4_PRSs_exacerbations_adj_copd, rownames = FALSE)
```

# TableS_PRSmix_assocsOutcomes: take mix_PRS for other outcomes, predict all other outcomes
## take mix_PRS for other outcomes, predict net exacerbations
```{r message=FALSE, warning=FALSE}
# without adjusting for fev1
mix_prs_list <- grep("^mix_prs_", names(df), value = TRUE)

results_mix_prss_exac <- data.frame()

for (prs in mix_prs_list) {
  model <- build_model(df, "total_Net_Exacerbations", prs, pcs, covars, model_type = "nb", adjustment="offset(log(total_days_since_P1/365))")
  beta <- coef(model)[prs]
  ci <- confint(model, prs, level = 0.95)
  beta_ci <- sprintf("%.3f (%.3f, %.3f)", beta, ci[1], ci[2])
  se <- summary(model)$coefficients[prs, "Std. Error"]
  pval <- summary(model)$coefficients[prs, 4]
  results_mix_prss_exac <- rbind(results_mix_prss_exac, 
                                 data.frame(prs = prs, beta = beta, ci_lower = ci[1], ci_upper = ci[2], 
                                            beta_ci = beta_ci, se = se, pval = pval))
}
results_mix_prss_exac <- results_mix_prss_exac %>%  mutate_if(is.numeric, signif, digits=3)
DT::datatable(results_mix_prss_exac, rownames = FALSE)
```

## take mix_PRS for other outcomes, predict copd
```{r message=FALSE, warning=FALSE}
results_mix_prss_copd <- data.frame()

for (prs in mix_prs_list) {
  model <- build_model(df, "copd", prs, pcs, covars, model_type = "binomial", adjustment=NULL)
  beta <- coef(model)[prs]
  ci <- confint(model, prs, level = 0.95)
  beta_ci <- sprintf("%.3f (%.3f, %.3f)", beta, ci[1], ci[2])
  se <- summary(model)$coefficients[prs, "Std. Error"]
  pval <- summary(model)$coefficients[prs, 4]
  results_mix_prss_copd <- rbind(results_mix_prss_copd, 
                                 data.frame(prs = prs, beta = beta, ci_lower = ci[1], ci_upper = ci[2], 
                                            beta_ci = beta_ci, se = se, pval = pval))
}
results_mix_prss_copd <- results_mix_prss_copd %>%  mutate_if(is.numeric, signif, digits=3)
DT::datatable(results_mix_prss_copd, rownames = FALSE)
```

## take mix_PRS for other outcomes, predict lung density
```{r}
results_mix_prss_ld <- data.frame()

for (prs in mix_prs_list) {
  model <- build_model(df, "lung_density_vnb_P1", prs, pcs, covars, model_type = "lm", adjustment = "scanner_model_clean_P1")
  beta <- coef(model)[prs]
  ci <- confint(model, prs, level = 0.95)
  beta_ci <- sprintf("%.3f (%.3f, %.3f)", beta, ci[1], ci[2])
  se <- summary(model)$coefficients[prs, "Std. Error"]
  pval <- summary(model)$coefficients[prs, 4]
  results_mix_prss_ld <- rbind(results_mix_prss_ld, 
                                 data.frame(prs = prs, beta = beta, ci_lower = ci[1], ci_upper = ci[2], 
                                            beta_ci = beta_ci, se = se, pval = pval))
}
results_mix_prss_ld <- results_mix_prss_ld %>%  mutate_if(is.numeric, signif, digits=3)
DT::datatable(results_mix_prss_ld, rownames = FALSE)
```

## take mix_PRS for other outcomes, predict Pi10
```{r}
results_mix_prss_pi10 <- data.frame()

for (prs in mix_prs_list) {
  model <- build_model(df, "Pi10_Thirona_P1", prs, pcs, covars, model_type = "lm", adjustment = "scanner_model_clean_P1")
  beta <- coef(model)[prs]
  ci <- confint(model, prs, level = 0.95)
  beta_ci <- sprintf("%.3f (%.3f, %.3f)", beta, ci[1], ci[2])
  se <- summary(model)$coefficients[prs, "Std. Error"]
  pval <- summary(model)$coefficients[prs, 4]
  results_mix_prss_pi10 <- rbind(results_mix_prss_pi10, 
                                 data.frame(prs = prs, beta = beta, ci_lower = ci[1], ci_upper = ci[2], 
                                            beta_ci = beta_ci, se = se, pval = pval))
}
results_mix_prss_pi10 <- results_mix_prss_pi10 %>%  mutate_if(is.numeric, signif, digits=3)
DT::datatable(results_mix_prss_pi10, rownames = FALSE)
```


```{r}
exac_results <- results_mix_prss_exac[, c("beta_ci", "pval")]
colnames(exac_results) <- c("exac_beta_ci", "exac_pval")

ld_results <- results_mix_prss_ld[, c("beta_ci", "pval")]
colnames(ld_results) <- c("lung_density_beta_ci", "lung_density_pval")

pi10_results <- results_mix_prss_pi10[, c("beta_ci", "pval")]
colnames(pi10_results) <- c("pi10_beta_ci", "pi10_pval")

copd_results <- results_mix_prss_copd[, c("beta_ci", "pval")]
colnames(copd_results) <- c("copd_beta_ci", "copd_pval")

TableS_PRSmix_assocsOutcomes <- cbind(exac_results, ld_results, pi10_results, copd_results)
write.table(TableS_PRSmix_assocsOutcomes, "data/TableS4_PRSmix_assocsOutcomes_cgNhw.txt", sep = "\t", row.names = TRUE, quote = FALSE)

DT::datatable(TableS_PRSmix_assocsOutcomes, options=list(autoWidth = TRUE, scrollX = '400px'))
```

# Supplemental figures
## Scatter plot of Pi10 and exacerbations
```{r fig.width=7, fig.height=5}
scatter_exac_pi10 <- ggplot(df, aes(x = exacerbation_rate, y = Pi10_Thirona_P1, color = copd)) +
  geom_point() +
  geom_smooth(method = "lm") + 
  # scale_x_log10() + 
  xlim(0, 5) +
  labs(x = "Annualized exacerbation rate", y = "Pi10", title = "Scatter plot of exacerbation vs. Pi10 stratified by COPD") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

scatter_exac_pi10
```

## Scatter plot of PRSmix and exacerbations
```{r fig.width=7, fig.height=5}
scatter_exac_copd <- ggplot(df, aes(x = exacerbation_rate, y = mix_prs_copd, color = copd)) +
  geom_point() +
  geom_smooth(method = "lm") + 
  # scale_x_log10() + 
  xlim(0, 5) +
  labs(x = "Annualized exacerbation rate", y = "PRSmix", title = "Scatter plot of exacerbation vs. PRSmix stratified by COPD") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

scatter_exac_copd
```

## ROC of COPD predicted by PRSmix vs FF PRS
```{r fig.width=7, fig.height=5, message=FALSE, warning=FALSE}
df_copd_complete <- df[complete.cases(df[, "copd"]), ]
outcome_binary <- "copd"

# print DeLong p-values for PRS-mix vs FF PRS 
copd_roc_out <- compare_roc_5_models(df_copd_complete, outcome_binary, mix_prs="mix_prs_copd", prs_to_compare="FF_ePRS", pcs, covars, adjustment=NULL, plot_title = "Outcome: COPD")

DT::datatable(copd_roc_out$auc_table)
DT::datatable(copd_roc_out$pval_table)

write.table(copd_roc_out$auc_table, "data/Table2_auc_copd_cgNhw.txt", sep = "\t", row.names = FALSE, quote = FALSE)
write.table(copd_roc_out$pval_table, "data/Table2_auc_pval_copd_cgNhw.txt", sep = "\t", row.names = FALSE, quote = FALSE)
```


```{r results='hide', message=FALSE, warning=FALSE}
pdf("plots/roc_copd_cgNhw.pdf", width = 7, height = 5)
compare_roc_5_models(df_copd_complete, outcome_binary, mix_prs="mix_prs_copd", prs_to_compare="FF_ePRS", pcs, covars, adjustment=NULL, plot_title = "Outcome: COPD")
invisible(dev.off())

jpeg("plots/roc_copd_cgNhw.jpg", width = 7, height = 5, units = "in", res = 600)
compare_roc_5_models(df_copd_complete, outcome_binary, mix_prs="mix_prs_copd", prs_to_compare="FF_ePRS", pcs, covars, adjustment=NULL, plot_title = "Outcome: COPD")
invisible(dev.off())
```

## ROC of exacerbations predicted by PRSmix vs FF PRS
```{r fig.width=7, fig.height=5, message=FALSE, warning=FALSE}
df_exac_complete <- df[complete.cases(df[, c("total_Net_Exacerbations", "total_days_since_P1")]), ]

outcome <- "total_Net_Exacerbations"
outcome_binary <- paste0(outcome, "_binary")
df_exac_complete[[outcome_binary]] <- ifelse(df_exac_complete[[outcome]] >= 2, 1, 0)

# print DeLong p-values for PRS-mix vs FF PRS
exac_roc_out <- compare_roc_5_models(df_exac_complete, outcome_binary, mix_prs="mix_prs_copd", prs_to_compare="FF_ePRS", pcs, covars, adjustment="offset(log(total_days_since_P1/365))", plot_title = "Outcome: frequent exacerbator")

DT::datatable(exac_roc_out$auc_table)
DT::datatable(exac_roc_out$pval_table)

write.table(exac_roc_out$auc_table, "data/Table2_auc_exac_cgNhw.txt", sep = "\t", row.names = FALSE, quote = FALSE)
write.table(exac_roc_out$pval_table, "data/Table2_auc_pval_exac_cgNhw.txt", sep = "\t", row.names = FALSE, quote = FALSE)
```

```{r results='hide', message=FALSE, warning=FALSE}
pdf("plots/roc_exac_cgNhw.pdf", width = 7, height = 5)
compare_roc_5_models(df_exac_complete, outcome_binary, mix_prs="mix_prs_copd", prs_to_compare="FF_ePRS", pcs, covars, adjustment="offset(log(total_days_since_P1/365))", plot_title = "Outcome: frequent exacerbator")
invisible(dev.off())

jpeg("plots/roc_exac_cgNhw.jpg", width = 7, height = 5, units = "in", res = 600)
compare_roc_5_models(df_exac_complete, outcome_binary, mix_prs="mix_prs_copd", prs_to_compare="FF_ePRS", pcs, covars, adjustment="offset(log(total_days_since_P1/365))", plot_title = "Outcome: frequent exacerbator")
invisible(dev.off())
```

# nominal effects
```{r message=FALSE, warning=FALSE}
nominal_effect_exac <- data.frame()

for (prs in prs_list) {
  model <- build_model(df, "total_Net_Exacerbations", prs, pcs, covars, model_type = "nb", adjustment="offset(log(total_days_since_P1/365))")
  beta <- coef(model)[prs]
  ci <- confint(model, prs, level = 0.95)
  beta_ci <- sprintf("%.3f (%.3f, %.3f)", beta, ci[1], ci[2])
  se <- summary(model)$coefficients[prs, "Std. Error"]
  pval <- summary(model)$coefficients[prs, 4]
  nominal_effect_exac <- rbind(nominal_effect_exac, 
                               data.frame(prs = prs, beta = beta, ci_lower = ci[1], ci_upper = ci[2], 
                                          beta_ci = beta_ci, se = se, pval = pval))
}

nominal_effect_exac <-nominal_effect_exac %>%  mutate_if(is.numeric, signif, digits=3)
DT::datatable(nominal_effect_exac, rownames = FALSE)
```

```{r message=FALSE, warning=FALSE}
nominal_effect_copd <- data.frame()

for (prs in prs_list) {
  model <- build_model(df, "copd", prs, pcs, covars, model_type = "binomial", adjustment=NULL)
  beta <- coef(model)[prs]
  ci <- confint(model, prs, level = 0.95)
  beta_ci <- sprintf("%.3f (%.3f, %.3f)", beta, ci[1], ci[2])
  se <- summary(model)$coefficients[prs, "Std. Error"]
  pval <- summary(model)$coefficients[prs, 4]
  nominal_effect_copd <- rbind(nominal_effect_copd, 
                               data.frame(prs = prs, beta = beta, ci_lower = ci[1], ci_upper = ci[2], 
                                          beta_ci = beta_ci, se = se, pval = pval))
}

nominal_effect_copd <-nominal_effect_copd %>%  mutate_if(is.numeric, signif, digits=3)
DT::datatable(nominal_effect_copd, rownames = FALSE)
```

```{r}
cor(df$BMI_P1, df$BMI_pgs)
```


# Heritability
```{r}
library(fmsb)

## logistic liability

## see appendix B from Lee et al. (https://core.ac.uk/reader/15145948?utm_source=linkout)

## LTRC

calcR2 <- function(dfm,pcsm,prs,ot,K){


  require(rms)

  # library(Design)

  require(pROC)



  ## Choose cohort for calculations

  # dfm <- pheno.ltrc

  # pcsm <- pcs.ltrc

  # ot="ipf.clinpath"

  # prs="prs_ipf"



  # define parameters

  dfm[,ot] <- as.numeric(dfm[,ot])

  nt=length(dfm[,1])

  ncase=sum(dfm[,ot]==1,na.rm=T)

  ncont=sum(dfm[,ot]==0,na.rm = T)

  # K=63/1e5 # population prevalence

  thd=-qnorm(K,0,1) # the threshold on the normal distribution which   truncates the proportion

  P=ncase/nt # proportion of cases in the case-control samples

  zv=dnorm(thd)

  mv=zv/K

  mv2=-mv*K/(1-K)


  

# probit

  pmv=glm(as.formula(paste0("as.numeric(",ot,")~",prs,"+",

                             paste0(pcsm,collapse="+"))),

          data=dfm,family=binomial(probit)) # probit model

  R2=var(pmv$linear.predictors)/(var(pmv$linear.predictors)+1) 

  print(paste0("R2 probit: ",signif(R2,3)))

  

  #logistic liability

  lrmv=glm(as.formula(paste0("as.numeric(",ot,")~",prs,"+",

                             paste0(pcsm,collapse="+"))),

          data=dfm,family=binomial(logit)) # logistic model

  R2.glm=var(lrmv$linear.predictors)/(var(lrmv$linear.predictors)+pi^2/3)

  print(paste0("R2 logistic liability: ",signif(R2.glm,3)))

  

  # R2 on the liability scale using AUC

  # aucv=auc(dfm[,ot],pmv$fitted.values)

  # qv=qnorm(aucv[1]) #Q in equation (11)

  # R2=2*qv^2/((mv2-mv)^2+qv^2*mv*(mv-thd)+mv2*(mv2-thd))

  # print(paste0("R2 liability using AUC: ",signif(R2,3)))

  

  # R2 on the liability scale using the transformation

  lmv=lm(as.formula(paste0("as.numeric(",ot,")~",prs,"+",

                             paste0(pcsm,collapse="+"))),

          data=dfm) #linear model

  R2O=var(lmv$fitted.values)/(ncase/nt*ncont/nt) #R2 on the observed scale

  theta=mv*(P-K)/(1-K)*(mv*(P-K)/(1-K)-thd) #θ in equation (15)

  cv=K*(1-K)/zv^2*K*(1-K)/(P*(1-P)) #C in equation (15)

  R2=R2O*cv/(1+R2O*theta*cv)


  print(paste0("R2 on liability scale using transformation: ",signif(R2,3)))




}
```

```{r}
calcR2(dfm=df, pcsm=pcs, prs="mix_prs_copd",ot="copd",K=0.5)
```

